\chapter{FAQ}
\label{sec:user_faq}

This chapter will give valuable advice on how to solve several tasks, solve
problems, and answer other questions related to the Visual Service Design Tool.


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{How to...}

\paragraph*{\dots draw Flow Objects on the Canvas?}
Although sometimes such diagrams can be seen, Flow Objects \emph{can not} be drawn
on the canvas.  Instead, each Flow Object has to be contained in a Lane, and the
Lane has to be inside of a Pool.  However, according to the specification one
Pool per diagram may have an invisible border, which can be set using the property
sheets.


\paragraph*{\dots create an Embedded Sub Process?}
First, create a basic Activity.  Then, in the Activity's property sheet, set the
activity type to \textsc{Embedded Sub Process}.  Now you can resize the Activity/Sub
Process to an appropriate size and create further Flow Objects inside of it.


\paragraph*{\dots create an Intermediate Event on an Activity's boundary?}
For creating an Intermediate Event on an Activity's boundary, i.e. an event- or
error handler, you should select the Event type from the palette and click the
Activity's label or boundary, but \emph{not} its compartment (in case it is a
subprocess).  Alternatively, you can use the miniature palette that appears when
hovering the mouse over the Activity's label or boundary.  Once created, the
Intermediate Event can be moved freely around the Activity's border.


\paragraph*{\dots draw Artifacts inside of a Pool?}
Contrary to Flow Objects, Artifacts can not be created inside of a Pool, but only
directly on the canvas.  However, you can create the Artifact on the canvas and
drag it over the Pool.  But remember that the Artifact is \emph{over}, and not
\emph{in}, the Pool, so it will not be moved together with the Pool.


\paragraph*{\dots make Assignments to Service Parameters?}
The arguments and results of service invocations are being stored in the Service's
input and output Properties.  So if a service is to be called with the argument
``Hello World'', that value has to be assigned to the service's respective input
Property using the \emph{Organize Assignments Dialog} of the Activity calling the
service.  Note, that Assignments to the input parameters have to be made
\emph{before}, and Assignments storing the output parameters in local variables
have to be made \emph{after} the Activities execution.  The Assign Time can be
specified in the Dialog, too.


A more convenient way, and sufficient in most situations, is to use the
\emph{Parameter Assignments Dialog}, which will take care of most of these details.
Of course, the same applies to Message Events and Send and Receive Activities, as
well.

%	\paragraph*{\dots enter a Time Date value?}
%	This value is expected in quite complex form, because it has to be processed by
%	a parser.  The value has to be in the form "yyyy-MM-dd'T'HH:mm:ss.SSSZ", e.g.
%	"2007-02-12T14:53:00.000+0100" for Monday, the 12th of February 2007 at 14:53:00
%	CET.


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Troubleshooting}

\paragraph*{My diagram file is broken. How can I fix it?}
In case the diagram is broken, it can be recreated from the model.  In case you
are using separate files for semantic model and notational model, delete the
diagram file \texttt{vsdt\_diagram} and create a new one by right-clicking the
model file \texttt{vsdt} and selecting the respective menu item.  In case you are
using one file for both semantic and notational model, \texttt{vsdtd}, open the
file in a text editor and remove the \texttt{notation:Diagram} elements (one for
each diagram contained in the file).  Now rename the file to \texttt{vsdt} and
initialize the diagram file (in this case, the newly generated file will contain
both the diagram and the model, so you should delete the old model file afterwards).
In both cases, you will have to recreate the diagram's layout.


\paragraph*{I've drawn a legal Business Process Diagram, but when I try to export
it the resulting code is broken.}
The diagram has to be in a certain form so it can successfully be transformed.
The diagram has to be \emph{structured}.  Primarily, there must not be blocks or
loops with multiple entry or exit points.  Although the VSDT can handle some forms
of slight unstructuredness, for best results you should design your diagram in a
way it can be mapped to a block-oriented language in a straight-forward manner.

If you are not sure whether the VSDT can handle all the structures used in the
diagram, use the \emph{Structure View} to check how the structure will look like
after the transformation.


\paragraph*{I've created a Business Process Diagram using element XYZ, but in the
resulting code this element is missing.}
Not each single feature of BPMN can be taken into account in all of the
transformations yet.  For those elements that can not be mapped, a no-operation
element will be created in the target model, such as a \verb_empty_ Activity in
BPEL or a \verb_logwarn_ in JIAC.  Be sure to substitute these elements with a
proper implementation after the transformation.

