\chapter{FAQ}
\label{sec:user_faq}

This chapter will give valuable advice on how to solve several tasks, solve
problems, and answer other questions related to the Visual Service Design Tool.


\section{How to...}
\label{sec:user_faq_howto}

\paragraph*{\dots draw Flow Objects on the Canvas?}
Although sometimes such diagrams can be seen, Flow Objects \emph{can not} be drawn
on the canvas.  Instead, each Flow Object has to be contained in a Lane, and the
Lane has to be inside of a Pool.  However, according to the specification one
Pool per diagram may have an invisible border, which can be set using the property
sheets.

\paragraph*{\dots create an Embedded Sub Process?}
First, create a basic Activity.  Then, in the Activity's property sheet, set the
activity type to \textsc{Embedded Sub Process}.  Now you can resize the Activity/Sub
Process to an appropriate size and create further Flow Objects inside of it.

\paragraph*{\dots create an Intermediate Event on an Activity's boundary?}
For creating an Intermediate Event on an Activity's boundary, i.e. an event- or
error handler, you should select the Event type from the palette and click the
Activity's label or boundary, and \emph{not} the compartment.  Alternatively, you
can use the miniature palette that appears when hovering the mouse over the
Activity's label or boundary.  Once created, the Intermediate Event can be moved
freely around the Activity's border.

\paragraph*{\dots draw Artifacts inside of a Pool?}
Contrary to Flow Objects, Artifacts can not be created inside of a Pool.  However,
you can create the Artifact on the canvas and drag it over the Pool.  But remember
that the Artifact is \emph{over}, and not \emph{in}, the Pool, so it will not be
moved together with the Pool.

\paragraph*{\dots make Assignments to Message Parameters?}
The arguments and results of service invocations are being stored in the Properties
of the input Message and the output Message.  So if a service is to be called
with the argument "Hello World", that value has to be assigned to the respective
Property of the service's input Message using the \emph{Organize Assignments
Dialog} of the Activity calling the service.  Note, that Assignments to the input
parameters have to be made \emph{before}, and Assignments storing the output
parameters in local variables have to be made \emph{after} the Activities
execution.  The Assign Time can be specified in the Dialog, too.  However, a more
convenient way, and sufficient in most situations, is to use the \emph{Parameter
Assignments Dialog}, which will take care of most of these details.  Of course,
the same applies to Message Events and Send and Receive Activities, as well.

%	\paragraph*{\dots enter a Time Date value?}
%	This value is expected in quite complex form, because it has to be processed by
%	a parser.  The value has to be in the form "yyyy-MM-dd'T'HH:mm:ss.SSSZ", e.g.
%	"2007-02-12T14:53:00.000+0100" for Monday, the 12th of February 2007 at 14:53:00
%	CET.


\section{Frequently Asked Questions}
\label{sec:user_faq_faq}

\paragraph*{My diagram is broken. How can I fix it?}
In case the diagram is broken, it can be recreated from the model.  In case you
are using separate files for semantic model and notational model, delete the
diagram file \texttt{vsdt_diagram} and create a new one by right-clicking the
model file \texttt{vsdt} and selecting the respective menu item.  In case you are
using one file for both semantic and notational model, \texttt{vsdtd}, open the
file in a text editor and remove the \texttt{notation:Diagram} elements (one for
each diagram contained in the file).  Now rename the file to \texttt{vsdt} and
initialize the diagram file (in this case you can delete the model file afterwards,
as the model has been copied to the diagram file).  In both cases, you will have
to recreate the diagram's layout.


\section{Troubleshooting}
\label{sec:user_faq_trouble}

\paragraph*{I've drawn a legal Business Process Diagram, but when I try to export
it the resulting code is broken.}
The diagram has to be in a certain form so it can successfully be transformed.
The diagram has to be \emph{structured}.  Primarily, there must not be blocks or
loops with multiple entry or exit points.  Although the VSDT can handle some forms
of slight unstructuredness, for best results you should design your diagram in a
way it can be mapped to a block-oriented language in a straight-forward manner.

\paragraph*{I've created a Business Process Diagram using element XYZ, but in the
resulting code this element is missing.}
Not each single feature of BPMN can be taken into account in all of the
transformations yet.  For those elements that can not be mapped, a no-operation
element will be created in the target model, such as a \verb|empty| Activity in
BPEL or a \verb|logwarn| in JIAC.  Be sure to substitute these elements with a
proper implementation after the transformation.

