\chapter{Preferences}
\label{sec:user_preferences}

This chapter will explain the several preferences that can be set for configuring the VSDT for your personal needs.

The preference pages can be accessed by navigating to \emph{Window $\rightarrow$ Preferences\dots} in the menu and selecting \emph{VSDT Diagram} from the list. The following sections will each explain one of the pages.


\section{Basic GMF Settings}
\label{sec:user_preferences_gmf}

These settings are for improving performance by deactivating certain features of the GMF runtime, or for setting the default colors and fonts for certain diagram elements.  As these features are self-explaining we will not go further into detail here.


\section{Other Settings}
\label{sec:user_preferences_other}

These settings are specific for the Visual Service Design Tool as a BPMN editor:

\begin{itemize}
	\item \textbf{Author} sets the default author for all new Business Process Diagrams created.
	\item \textbf{Enable Activity Icons} activates additional markers placed in the corners of Activities. These markers use intuitive symbols to indicate the Activity's type and whether the Activity has any Assignments and/or Properties.
	\item \textbf{Use Additional Colors} helps to distinguish the several diagram elements by the use of colors.
\end{itemize}

\emph{Note} that both the Enable Activity Icons and the Use Additional Colors settings take effect only after closing and reopening a given Business Process Diagram file.
